---
title: "Patagona PSMC"
author: "Jessie Williamson"
date: "2/11/2022; last revised 2024-04-10"
output: html_document
---

PSMC is a little tricky to work with because the command line plotting function is a little rigid in terms of adjusting aesthetics, etc. 


This script is adapted from Jager et al. 2021, Scientific Reports. Their paper is here: https://www.nature.com/articles/s41598-021-83823-8.epdf?sharing_token=GbrMcCfB_r8Wl_s5OErRStRgN0jAjWel9jnR3ZoTv0MimupHo1Z8n-Xm9ErGcLGzooYFx6795KsWPLN1dt84JC86QWl1tKPE-7nl2rwKpiDDI4L6H-QkXd34Y33VnvtJocJr7-MRlPEoZhSvbXeTy2iLVRJZequvevVoB9daSXQ%3D
(And they have original R function they used)

And Jager et al. PSMC R code here: https://github.com/DeondeJager/Buffalo_PopGenomics/blob/master/03-Figures/22-Fig4_PSMC.R#L4

Script adapted for this data set from Emily Humble's ("3.1_psmc.R" from https://github.com/elhumble/SHO_analysis_2020)
Calls the "psmc.result" function from the "plot_psmc.R" script from https://figshare.com/articles/Plot_PSMC_results/3996156/1


# Clear workspace and set WD
```{R}
rm(list=ls(all=TRUE)) # clear workspace 
setwd("/Users/Jessie/MSBbirds Dropbox/Jessie Williamson/Rdirectory/Patagona")
```

# Load packages
```{r}
library(ggplot2)
source("1_r_scripts/psmc_plot_function.R") # plot_psmc function. 
library(data.table)
library(plyr)
library(tidyr)
library(dplyr)
library(scales)
options(scipen=999) # Disable scientific notation of large numbers- see "Scipen" under ?options
```


_____


# load in data 
```{R}
# Load psmc files (from psmc_to_plot folder because there are _run6.psmc files from other birds that we don't want to use)
# MAKE SURE YOU HAVE A FINAL FORWARD SLASH AT END OF FOLDER FILE PATH NAME!!!!
psmc_files <- paste("Genomics/psmc/psmc_to_plot/", list.files(path = "Genomics/psmc/psmc_to_plot/", pattern="*run6.psmc"), sep = "") 

# Import bootstrap files
# Need to download individual bootstrap files from CARC to local folder
# In the buffalo tutorial, they used *round* because of the naming convention of their PSMC bootstraps
boot_files <- paste("Genomics/psmc_boot_run6/", list.files(path = "Genomics/psmc_boot_run6/", pattern="*r*"), sep = "")
```


# Specify variables for plotting
Pay attention to these! 
```{r}
# Number of iterations to use in file
i.iteration=20 

# Bin size 
#  default is 100 for PSMC
s=100 

# Set mu and g (for plot_psmc function):
# Zebra finch mutation rate is 2.21x10^-9 (2.21e-9) (Nam et al. 2010)
# SUPER IMPORTANT! Remember that mu=mutation rate*generation time*10^-9
# Our value is 2.21*10^-9)*2.735 = 0.00000000604435
#mu <- 6.04435e-9 # see above! 
mu <- (2.21e-9)*2.735 # see above! 
g <- 2.735 # From Bird et al. 2020
```


# Run psmc function on .psmc and bootstrap files; wrangle and transform as needed
```{r}
# Run "psmc.result" from the "plot_psmc" function on psmc files 
psmc_pgig <- lapply(psmc_files, psmc.result, i.iteration=i.iteration, mu=mu, s=s, g=g)

## Set names in psmc_pgig to sample names
# Make sure you have final forward slash at the end of file path names! 
dataset_names <-list.files(path = "Genomics/psmc/psmc_to_plot/", pattern="*run6.psmc")
dataset_names <-  gsub(".psmc", "", dataset_names)
names(psmc_pgig) <- dataset_names

# Transform list into dataframe
psmc_pgig <- ldply(psmc_pgig, .id = "Sample")  

# Run "psmc.result" from the "plot_psmc" function on bootstrap files 
boot <- lapply(boot_files, psmc.result, i.iteration=i.iteration, mu=mu, s=s, g=g)

# Set names in psmc_pgig to sample names
dataset_names <-list.files(path = "Genomics/psmc_boot_run6/", pattern="*r*") # just want run naming pattern!
dataset_names <-  gsub(".psmc", "", dataset_names)
names(boot) <- dataset_names

# Transform list into dataframe
boot_df <- ldply(boot, .id = "ID") %>%
  separate(ID, c("Sample", "Boot"), sep = "_bootstrap_r*", remove = F) %>% # sep needs to correspond to my run naming convention
  mutate(Boot = gsub("_", "", Boot))

# Scale Ne (so that y-axis is in thousands vs x10^4)
boot_df$Ne_scaled <- boot_df$Ne/100000
psmc_pgig$Ne_scaled <- psmc_pgig$Ne/100000
#plot(boot_df$Ne, boot_df$YearsAgo)

# Drop YearsAgo <100,000
# I tried this to abridge x-axis...didn't work. 
#boot_sub <- boot_df[-which(boot_df$YearsAgo < 1000000 | morphosub$tail > 92),] # # 1 outlier
```


# Plotting all birds together 

# PSMC plot code
```{r}
#cbPalette <- c("#0066FF", "olivedrab3", "black", "#FF3300")
cbPalette <- c("#F2C900","#F2C900","#99B5BC","#99B5BC","#196C91","#196C91","#FF0000","#FF0000")
# Oorder of countries (not levels) goes: Argentina, Peru, Ecuador, Chile

# Argentina (mustard): #F2C900
# Peru (light dull blue): #99B5BC
# Ecuador (navy): #196C91
# Chile (red): #FF0000

# CRG script starts here: 
psmc_pgig$ID <- "dummy_ID" # Adding dummy ID column to make rbinding to boot_df possible
psmc_pgig$Boot <- 0 #likewise adding boot columns for rbinding

order <- names(boot_df) #Make vector of names for ordering
psmc_pgig <- psmc_pgig[, order] #reorder psmc_pgig

# ready to append to boot_df for plotting
psmc_pgig_comb <- rbind(psmc_pgig, boot_df)
year_trunc <- 9000 # set some value to truncate years... Play with this as you see fit.
  # Basically: recent year estimates are wonky, as expected; likely contribute to axis issues
  # The most recent you can set this is 9,000 years ago; less than that and the whole plot flattens out
  # Keep this truncate value to  cut off bad stuff and make sure plot looks good 

# PLOT - SIZED FOR FIGURE 2 PANEL IN PAPER
# 5.5 height and 4.5 width are dimensions we want, but fig is overall too small
pdf(file="./Genomics/psmc/PSMC_plot_AllBirds_ForFig2Panel_Width5.5Height5.5_2023-04-04.pdf", height=5.5, width=5.5)
ggplot(psmc_pgig_comb %>% filter(YearsAgo >= year_trunc, !ID %in% "dummy_ID"),
  mapping=aes(x=YearsAgo, y=Ne, color=Sample, group=factor(Sample)))+ # selecting raw #bootstraps by group to plot thin lines 
  geom_path(alpha=0.1, size=0.3)+ #controls thin lines 
  geom_line(psmc_pgig_comb%>%filter(YearsAgo >= year_trunc, ID %in% "dummy_ID"), 
            mapping=aes(x=YearsAgo, y=Ne, color=Sample, group=factor(Sample)), size=0.7)+ # Controls mean bootstrap values (thick lines)
  theme_classic() +
  scale_colour_manual(values = cbPalette,
                      labels = c("Argentina", "Arg Boot","Peru", "Peru Boot","Ecuador", "Ec Boot","Chile", "ChileBoot"),
                      name = expression(paste(bold("Population")))) +
  theme(legend.position = c(0.7,0.8)) + # order of your factors was wrong. This should be correct but look at data frame to make sure they're correct
  scale_x_log10(label = comma, # EFG thinks log transformation necessary bc mostly long windows so looks weird w/out it
                breaks = c(15000, 100000,1000000), limits=c(10000, 1800000)) + # adjust limits to cut off large values
               # ) + # Set x-axis limits, abridge present because estimates wonky limits=c(19000,10000000),expand = c(0,0)
  annotation_logticks(sides = "b", alpha = 0.5) +
   #coord_cartesian(ylim = c(0, 200000)) + 
  coord_cartesian(xlim = c(16000, 1200000)) + # This is important for constraining the axis to cut of wonky present and past values 
  #scale_y_continuous(label = comma) +
  scale_y_continuous(label = comma, breaks=c(0, 50000, 100000, 150000, 200000), limits=c(0, 200000)) +
  xlab(expression(paste("Years before present (g = 2.735, ", mu," = 6.04435e-9",")"))) +  
#  ylab(expression(paste("Effective Population Size",~italic(N[e]),x10^4))) + # If plotting Ne_scaled
  ylab(expression(paste("Effective Population Size",~italic(N[e])))) + # If NOT plotting Ne scaled
  geom_vline(xintercept=22000, linetype = "dashed", colour = "grey") + # Last Glacial Maximum
#  geom_vline(xintercept=180000, linetype = "dashed", colour = "grey") + # TRYING TO EYEBALL HISTORICAL DIVERGENCE 
    # Uncomment the line above if you want a dashed vertical line at historical divergence
  # #  geom_vline(xintercept = 193000, linetype = "dashed", colour = "grey") + # Subspecies split
  annotate("text", x = 25000, y=160000, label = "Last Glacial Maximum (LGM)", angle = -90, size = 4, colour="grey")
dev.off()
```

---


# Print environment for reproducibility
```{r}
sessionInfo() # List of packages and versions in use 
```

###########

## END 
